---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# modelselection

<!-- badges: start -->
<!-- badges: end -->

The goal of `modelselection` is to provide common model selection and tuning 
utilities in an intuitive manner. Specifically, I want something that is 
lightweight (so not `mlr3`) and doesn't force you to adopt a whole new modeling
paradigm (so not `tidymodels`). Also, I want the provided functionality to be
very type agnostic and be able to work with data frames, standard dense
matrices, and `Matrix` sparse matrices. Finally, I want it to be easily
distributable (It's built on top of the `future.apply` package) and I want to 
have full control over it 😉.

## Installation

You can install the development version of `modelselection` with:

``` r
# install.packages("devtools")
devtools::install_github("dmolitor/modelselection")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(modelselection)
## basic example code
```

